---
title: "CHIP Gene and VAF Subgroup Survival Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Downloads/CHIP:HCC project/7. Correlative Analysis")
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.width = 7, fig.height = 6)
```

# Introduction

This report classifies each patient's dominant CHIP clone by **gene** (TET2,
DNMT3A, or Other) and by **VAF level** (High, >10%; Low, <10%), then groups
patients into four categories:

- **CHIP-**: no CHIP
- **TET2_High**: dominant clone is TET2 with VAF > 10%
- **Other_High**: dominant clone is non-DNMT3A/non-TET2 with VAF > 10%
- **Other_CHIP**: all remaining CHIP+ patients (every low-VAF case, plus all
  DNMT3A cases regardless of VAF)

It then evaluates whether these groups are associated with overall survival
(OS), progression-free survival (PFS), time to progression (TTP), and
fibrosis level. Each analysis is run once on all patients, then repeated
separately for resection and transplant patients.

# Data Preparation

## Load clinical data

```{r load-data}
library(dplyr)
library(stringr)
library(survival)
library(lubridate)
library(ggsurvfit)
library(ggplot2)

dir.create("Plots/GeneVAF_Subgroups", showWarnings = FALSE, recursive = TRUE)

clinical <- read.csv("clinical_final.csv")

# Most dates are entered as YYYY-MM-DD, but a few rows use MM-DD-YYYY. Try ymd()
# first and fall back to mdy() for anything that fails to parse, instead of
# silently dropping those patients from the time-to-event outcomes.
parse_date_flex <- function(x) as_date(parse_date_time(x, orders = c("ymd", "mdy"), quiet = TRUE))
```

## Classify each patient's dominant CHIP clone

Each patient's `CHIP_vaf` field is formatted as `GENE_vaf||GENE_vaf...`. For
patients with more than one CHIP mutation, the clone with the highest VAF is
taken as the dominant clone; its gene and VAF determine that patient's group.

```{r parse-clone}
split_underscore <- function(x, index) strsplit(x, "_", fixed = TRUE)[[1]][index]

parse_dominant_clone <- function(x) {
  if (is.na(x) || nchar(x) == 0) {
    return(data.frame(gene = NA_character_, vaf = 0))
  }
  parts <- strsplit(x, "||", fixed = TRUE)[[1]]
  genes <- unlist(lapply(parts, split_underscore, index = 1))
  vafs <- as.numeric(unlist(lapply(parts, split_underscore, index = 2)))
  data.frame(gene = genes[which.max(vafs)], vaf = max(vafs))
}

dominant <- do.call(rbind, lapply(clinical$CHIP_vaf, parse_dominant_clone))
clinical$dominant_gene <- dominant$gene
clinical$max_vaf <- dominant$vaf

classify_gene <- function(gene) {
  ifelse(gene == "TET2", "TET2", ifelse(gene == "DNMT3A", "DNMT3A", "Other"))
}

# Like parse_dominant_clone(), but returns every clone for a patient instead
# of just the dominant one.
parse_all_clones <- function(x) {
  if (is.na(x) || nchar(x) == 0) return(data.frame(gene = character(0), vaf = numeric(0)))
  parts <- strsplit(x, "||", fixed = TRUE)[[1]]
  genes <- unlist(lapply(parts, split_underscore, index = 1))
  vafs <- as.numeric(unlist(lapply(parts, split_underscore, index = 2)))
  data.frame(gene = genes, vaf = vafs)
}
```

## Define VAF and gene subgroups

```{r define-groups}
# 1. VAF split: high (>10%) vs low (<10%)
clinical$vaf_group <- ifelse(clinical$CHIP == 0, "CHIP-",
                       ifelse(clinical$max_vaf > 0.10, "High", "Low"))

# 2. Gene category: TET2, DNMT3A, or Other
clinical$gene_category <- ifelse(clinical$CHIP == 0, "None",
                           ifelse(clinical$dominant_gene == "TET2", "TET2",
                           ifelse(clinical$dominant_gene == "DNMT3A", "DNMT3A", "Other")))

# 3. Combined group:
#    - CHIP-: no CHIP
#    - TET2_High: dominant clone is TET2 with VAF > 10%
#    - Other_High: dominant clone is non-DNMT3A/non-TET2 (i.e. "Other") with VAF > 10%
#    - Other_CHIP: all remaining CHIP+ patients (all low VAF, plus all DNMT3A
#      regardless of VAF)
clinical$chip_group <- ifelse(clinical$CHIP == 0, "CHIP-",
                        ifelse(clinical$gene_category == "TET2" & clinical$vaf_group == "High", "TET2_High",
                        ifelse(clinical$gene_category == "Other" & clinical$vaf_group == "High", "Other_High",
                        "Other_CHIP")))

group_levels <- c("CHIP-", "TET2_High", "Other_High", "Other_CHIP")
clinical$chip_group <- factor(clinical$chip_group, levels = group_levels)
```

Patient counts per CHIP gene/VAF group:

```{r group-counts}
table(clinical$chip_group)
```

Patient counts per group, by surgery type (0 = resection, 1 = transplant):

```{r group-counts-by-surgery}
table(clinical$chip_group, clinical$type_sx)
```

### Multi-Mutation Patient Detail

`chip_group` above is decided solely by each patient's single highest-VAF
("dominant") clone. For every patient with more than one CHIP mutation, this
lists all of their mutations and VAFs, ordered by patient ID, along with the
group **each individual clone** would produce on its own (the same gene x
VAF rule as above). When those per-clone labels disagree (flagged by
`ambiguous`), the patient could have been classified into more than one
combined group depending on which clone is treated as dominant - this can
happen either because the clones are in different genes, or because the
same gene straddles the 10% VAF threshold.

```{r multi-mutation-detail}
clone_would_be_group <- function(gene, vaf) {
  category <- classify_gene(gene)
  vaf_grp <- ifelse(vaf > 0.10, "High", "Low")
  ifelse(category == "TET2" & vaf_grp == "High", "TET2_High",
  ifelse(category == "Other" & vaf_grp == "High", "Other_High", "Other_CHIP"))
}

multi_mutation_detail <- do.call(rbind, lapply(seq_len(nrow(clinical)), function(i) {
  clones <- parse_all_clones(clinical$CHIP_vaf[i])
  if (nrow(clones) < 2) return(NULL)
  clones$would_be_group <- clone_would_be_group(clones$gene, clones$vaf)
  data.frame(study_id = clinical$study_id[i], gene = clones$gene, vaf = clones$vaf,
             would_be_group = clones$would_be_group,
             ambiguous = length(unique(clones$would_be_group)) > 1,
             assigned_group = as.character(clinical$chip_group[i]))
}))
multi_mutation_detail <- multi_mutation_detail[order(multi_mutation_detail$study_id,
                                                      -multi_mutation_detail$vaf), ]
multi_mutation_detail
```

## Alternate CHIP Grouping

A second, simpler 3-group classification, used later for a separate set of
results (see the "Results (Alternate CHIP Grouping)" section):

- **CHIP-**: no CHIP
- **NonDNMT3A_High**: dominant clone is any gene other than DNMT3A with VAF > 10%
- **Other_CHIP**: all remaining CHIP+ patients (DNMT3A regardless of VAF,
  plus all low-VAF cases)

```{r define-alt-groups}
clinical$chip_group_alt <- ifelse(clinical$CHIP == 0, "CHIP-",
                            ifelse(clinical$gene_category != "DNMT3A" & clinical$vaf_group == "High",
                                   "NonDNMT3A_High", "Other_CHIP"))

alt_group_levels <- c("CHIP-", "NonDNMT3A_High", "Other_CHIP")
clinical$chip_group_alt <- factor(clinical$chip_group_alt, levels = alt_group_levels)
```

Patient counts per alternate CHIP group:

```{r alt-group-counts}
table(clinical$chip_group_alt)
```

Patient counts per alternate group, by surgery type (0 = resection, 1 = transplant):

```{r alt-group-counts-by-surgery}
table(clinical$chip_group_alt, clinical$type_sx)
```

## Patients with Multiple CHIP Calls

Some patients carry more than one CHIP mutation at once (e.g. a TET2 clone
alongside a DNMT3A or other-gene clone). Every group assignment above is
based solely on that patient's single highest-VAF ("dominant") clone -
any other co-occurring mutations are not otherwise reflected in `chip_group`.
This section characterizes those multi-mutation patients and shows where
the dominant-clone rule placed them.

```{r multi-chip-parse}
multi_summary <- do.call(rbind, lapply(seq_len(nrow(clinical)), function(i) {
  clones <- parse_all_clones(clinical$CHIP_vaf[i])
  if (nrow(clones) == 0) {
    return(data.frame(study_id = clinical$study_id[i], n_mutations = 0,
                       n_genes = 0, n_categories = 0, categories = NA_character_))
  }
  clones$category <- classify_gene(clones$gene)
  data.frame(study_id = clinical$study_id[i], n_mutations = nrow(clones),
             n_genes = length(unique(clones$gene)),
             n_categories = length(unique(clones$category)),
             categories = paste(sort(unique(clones$category)), collapse = "+"))
}))

clinical <- left_join(clinical, multi_summary, by = "study_id")

# Multi-mutation: more than one CHIP clone detected.
# Multi-category: those clones span more than one of TET2 / DNMT3A / Other -
# this is the "both TET2 and non-TET2" scenario referenced above.
clinical$multi_mutation <- clinical$n_mutations > 1
clinical$multi_category <- clinical$n_categories > 1
```

Number of CHIP mutations per patient (CHIP+ only):

```{r multi-chip-mutation-count}
table(clinical$n_mutations[clinical$CHIP == 1])
```

Of patients with more than one CHIP mutation, how many span more than one
gene category:

```{r multi-chip-category-count}
table(clinical$categories[clinical$multi_mutation])
```

Final `chip_group` assigned to multi-category patients under the
dominant-clone rule:

```{r multi-chip-final-group}
table(clinical$chip_group[clinical$multi_category])
```

```{r multi-chip-plots}
# Plot 1: how many CHIP mutations does each CHIP+ patient carry.
plot_mutation_count <- ggplot(filter(clinical, CHIP == 1),
                               aes(x = factor(n_mutations))) +
  geom_bar() +
  labs(x = "Number of CHIP mutations detected", y = "Number of patients",
       title = "CHIP mutation burden per patient") +
  theme_classic()
ggsave("Plots/GeneVAF_Subgroups/MultiCHIP_MutationCount.pdf",
       plot = plot_mutation_count, height = 5, width = 6)
print(plot_mutation_count)

# Plot 2: among multi-mutation patients, do their clones stay within one gene
# category or span multiple (e.g. TET2 + Other)?
plot_category_breakdown <- ggplot(filter(clinical, multi_mutation),
                                   aes(x = categories)) +
  geom_bar() +
  labs(x = "Gene categories represented among a patient's CHIP clones",
       y = "Number of patients",
       title = "Multi-mutation patients: single vs. multiple gene categories") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
ggsave("Plots/GeneVAF_Subgroups/MultiCHIP_CategoryBreakdown.pdf",
       plot = plot_category_breakdown, height = 5, width = 6)
print(plot_category_breakdown)

# Plot 3: for the multi-category patients specifically, which final chip_group
# did the dominant-clone rule assign them to?
plot_multicat_assignment <- ggplot(filter(clinical, multi_category),
                                    aes(x = droplevels(chip_group))) +
  geom_bar() +
  labs(x = "Final assigned group", y = "Number of patients",
       title = "Multi-category patients: final group after dominant-clone rule") +
  theme_classic()
ggsave("Plots/GeneVAF_Subgroups/MultiCHIP_FinalGroupAssignment.pdf",
       plot = plot_multicat_assignment, height = 5, width = 6)
print(plot_multicat_assignment)
```

# Analysis Helper Functions

The functions below are reused across every outcome and cohort in the
Results section. `run_km_cox()` fits a Cox model and draws/saves a
Kaplan-Meier plot for a time-to-event outcome, labeling each group with its
sample size and showing the model's overall logrank p-value as a subtitle.
`run_fibrosis()` cross-tabs fibrosis score (levels 0-4, taken directly from
the clinical data) against group, runs a chi-square test, and draws/saves a
bar plot of the fibrosis score distribution within each group, with each
group's sample size shown on the x-axis.

```{r helpers}
# Refit the group factor to only the levels present in a subset, keeping
# "CHIP-" as the reference level whenever it is present.
refactor_group <- function(df, group_col = "chip_group") {
  df[[group_col]] <- droplevels(df[[group_col]])
  if ("CHIP-" %in% levels(df[[group_col]])) {
    df[[group_col]] <- relevel(df[[group_col]], ref = "CHIP-")
  }
  df
}

run_km_cox <- function(df, time_col, event_col, group_col, ylab, label, outfile) {
  df <- refactor_group(df, group_col)
  n_levels <- nlevels(df[[group_col]])
  cat("n =", nrow(df), " groups present:", paste(levels(df[[group_col]]), collapse = ", "), "\n")

  if (n_levels < 2) {
    cat("Skipped: fewer than 2 groups present.\n")
    return(invisible(NULL))
  }

  form <- as.formula(paste0("Surv(", time_col, ", ", event_col, ") ~ ", group_col))

  cox_res <- tryCatch(coxph(form, data = df), error = function(e) {
    cat("Cox model failed:", conditionMessage(e), "\n")
    NULL
  })
  if (!is.null(cox_res)) print(summary(cox_res))

  p_val <- if (!is.null(cox_res)) {
    tryCatch(summary(cox_res)$sctest[["pvalue"]], error = function(e) NA)
  } else NA
  p_label <- if (is.na(p_val)) "NA" else format.pval(p_val, digits = 3, eps = 0.001)

  # Relabel groups with sample size for the plot only (leaves the Cox model
  # term names in the summary above untouched).
  plot_df <- df
  group_counts <- table(plot_df[[group_col]])
  levels(plot_df[[group_col]]) <- paste0(levels(plot_df[[group_col]]), " (n=", as.integer(group_counts), ")")

  plot <- tryCatch({
    survfit2(form, data = plot_df) %>%
      ggsurvfit() +
      labs(x = "Days", y = ylab, title = label,
           subtitle = paste0("Logrank p = ", p_label)) +
      ylim(c(0, 1.0))
  }, error = function(e) {
    cat("KM plot failed:", conditionMessage(e), "\n")
    NULL
  })

  if (!is.null(plot)) {
    ggsave(outfile, plot = plot, height = 6, width = 7)
    print(plot)
  }

  invisible(cox_res)
}

run_fibrosis <- function(df, group_col, label, table_outfile, plot_outfile) {
  df <- refactor_group(df, group_col)
  n_levels <- nlevels(df[[group_col]])
  cat("n =", nrow(df), " groups present:", paste(levels(df[[group_col]]), collapse = ", "), "\n")

  if (n_levels < 2) {
    cat("Skipped: fewer than 2 groups present.\n")
    return(invisible(NULL))
  }

  tab <- table(df[[group_col]], df$fibrosis_score)
  colnames(tab) <- paste0("F", colnames(tab))
  print(tab)
  write.csv(tab, file = table_outfile)

  chisq_res <- tryCatch(
    chisq.test(tab, simulate.p.value = TRUE, B = 10000),
    error = function(e) {
      cat("Chi-square test failed:", conditionMessage(e), "\n")
      NULL
    }
  )
  if (!is.null(chisq_res)) print(chisq_res)

  group_counts <- table(df[[group_col]])
  x_labels <- setNames(paste0(names(group_counts), "\n(n=", as.integer(group_counts), ")"),
                       names(group_counts))

  plot <- tryCatch({
    ggplot(df, aes(x = .data[[group_col]], fill = factor(fibrosis_score))) +
      geom_bar(position = "fill") +
      scale_x_discrete(labels = x_labels) +
      labs(x = NULL, y = "Proportion of patients", fill = "Fibrosis score",
           title = label) +
      theme_classic()
  }, error = function(e) {
    cat("Fibrosis plot failed:", conditionMessage(e), "\n")
    NULL
  })

  if (!is.null(plot)) {
    ggsave(plot_outfile, plot = plot, height = 5, width = 7)
    print(plot)
  }

  invisible(list(chisq = chisq_res))
}
```

# Preoperative Lab Values by CHIP Group

`clinical_data.csv` (not `clinical_final.csv`) holds the preoperative blood
work. Its `study_id` is formatted `HCC-ctDNA-NNN`; that prefix is stripped to
match the numeric `study_id` used elsewhere in this report. Several lab
columns were entered as free text (e.g. `"nr"`, `"not reported"`, or a value
with a parenthetical date like `"2.0 (19/01/2018)"`) - `parse_numeric_flex()`
pulls out the leading numeric token and treats anything else as missing.

```{r lab-data-prep}
parse_numeric_flex <- function(x) as.numeric(str_extract(trimws(as.character(x)), "^[0-9]+\\.?[0-9]*"))

lab_vars <- c(
  preop_bld_1_inr = "INR",
  preop_bld_1_bili = "Bilirubin",
  preop_bld_1_albumin = "Albumin",
  preop_bld_1_platelet = "Platelet count",
  preop_bld_1_creatinine = "Creatinine",
  preop_bld_1_alt = "ALT",
  preop_bld_1_ast = "AST",
  preop_bld_1_afp = "AFP",
  preop_bld_1_wbc = "WBC",
  preop_bld_1_hb = "Hemoglobin"
)

lab_data <- read.csv("clinical_data.csv")
lab_data$study_id <- as.numeric(gsub("HCC-ctDNA-", "", lab_data$study_id))
lab_data[names(lab_vars)] <- lapply(lab_data[names(lab_vars)], parse_numeric_flex)
lab_data <- select(lab_data, study_id, all_of(names(lab_vars)))
lab_data <- inner_join(lab_data, select(clinical, study_id, chip_group, chip_group_alt), by = "study_id")
```

`run_lab_distribution()` compares one lab value's distribution across CHIP
groups with a boxplot (with individual patients jittered on top) and a
Kruskal-Wallis test - a non-parametric test that suits the skewed,
outlier-heavy nature of lab values better than an ANOVA would. Each group is
labeled with its sample size.

```{r lab-distribution-helper}
run_lab_distribution <- function(df, var, var_label, group_col, plot_outfile) {
  df <- refactor_group(df, group_col)
  df <- df[!is.na(df[[var]]), ]
  df[[group_col]] <- droplevels(df[[group_col]])
  n_levels <- nlevels(df[[group_col]])
  cat("n =", nrow(df), " groups present:", paste(levels(df[[group_col]]), collapse = ", "), "\n")

  summary_tab <- df %>%
    group_by(.data[[group_col]]) %>%
    summarise(n = n(), median = median(.data[[var]]), mean = mean(.data[[var]]),
              sd = sd(.data[[var]]), q1 = quantile(.data[[var]], 0.25),
              q3 = quantile(.data[[var]], 0.75), .groups = "drop")
  print(summary_tab)

  if (n_levels < 2) {
    cat("Skipped: fewer than 2 groups present.\n")
    return(invisible(NULL))
  }

  kw_res <- tryCatch(kruskal.test(df[[var]], df[[group_col]]), error = function(e) {
    cat("Kruskal-Wallis test failed:", conditionMessage(e), "\n")
    NULL
  })
  if (!is.null(kw_res)) print(kw_res)
  p_label <- if (is.null(kw_res)) "NA" else format.pval(kw_res$p.value, digits = 3, eps = 0.001)

  group_counts <- table(df[[group_col]])
  x_labels <- setNames(paste0(names(group_counts), "\n(n=", as.integer(group_counts), ")"),
                       names(group_counts))

  plot <- tryCatch({
    ggplot(df, aes(x = .data[[group_col]], y = .data[[var]])) +
      geom_boxplot(outlier.shape = NA) +
      geom_jitter(width = 0.15, alpha = 0.4) +
      scale_x_discrete(labels = x_labels) +
      labs(x = NULL, y = var_label, title = var_label,
           subtitle = paste0("Kruskal-Wallis p = ", p_label)) +
      theme_classic()
  }, error = function(e) {
    cat("Lab distribution plot failed:", conditionMessage(e), "\n")
    NULL
  })

  if (!is.null(plot)) {
    if (var ==  'preop_bld_1_afp') {
          plot = plot + ylim(c(0,800))
    }
    else if (max(df[[var]] > 500)) { 
      plot = plot + ylim(c(0,200))
    } else if (var ==  'preop_bld_1_wbc') {
      plot = plot + ylim(c(0,20))
    }
    ggsave(plot_outfile, plot = plot, height = 5, width = 7)
    print(plot)
  }

  invisible(list(kruskal = kw_res))
}
```

The same lab distributions are produced twice: once with the original 4-group
classification (`chip_group`) and once with the alternate 3-group
classification (`chip_group_alt`). Original-grouping plots are written to
`Plots/GeneVAF_Subgroups/`; alternate-grouping plots to the `AltGrouping`
subfolder, leaving the original output files untouched.

```{r lab-distribution-run}
dir.create("Plots/GeneVAF_Subgroups/AltGrouping", showWarnings = FALSE, recursive = TRUE)

lab_groupings <- list(
  list(group_col = "chip_group",     out_dir = "Plots/GeneVAF_Subgroups",            label = "Original grouping"),
  list(group_col = "chip_group_alt", out_dir = "Plots/GeneVAF_Subgroups/AltGrouping", label = "Alternate grouping")
)

for (grouping in lab_groupings) {
  cat("\n========", grouping$label, "========\n")
  for (lab_col in names(lab_vars)) {
    cat("\n----", lab_vars[[lab_col]], "----\n")
    run_lab_distribution(lab_data, lab_col, lab_vars[[lab_col]], grouping$group_col,
                          plot_outfile = paste0(grouping$out_dir, "/Labs_", lab_col, ".pdf"))
  }
}
```

# Outcome Dataset Construction

```{r build-outcomes}
# ---- Time to Progression (TTP): surgery date -> recurrence, censored at last f/u ----
build_TTP <- function(clinical, group_col = "chip_group") {
  TTP <- select(clinical, study_id, chip_group = all_of(group_col), type_sx,
                date_sx, recurrence, status, date_recur_1, date_of_death, date_last_fu)

  TTP$recurrence[TTP$recurrence == 0] <- 2
  TTP$recurrence[TTP$recurrence == 1] <- 0
  TTP$recurrence[TTP$recurrence == 2] <- 1

  TTP$status[TTP$status == 0] <- 2
  TTP$status[TTP$status == 1] <- 0
  TTP$status[TTP$status == 2] <- 1

  TTP <- TTP %>%
    mutate(date_sx = parse_date_flex(date_sx),
           date_recur_1 = parse_date_flex(date_recur_1),
           date_of_death = parse_date_flex(date_of_death),
           date_last_fu = parse_date_flex(date_last_fu))

  TTP$date_last_fu[TTP$recurrence == 0 & TTP$status == 1 & is.na(TTP$date_last_fu)] <-
    TTP$date_of_death[TTP$recurrence == 0 & TTP$status == 1 & is.na(TTP$date_last_fu)]

  TTP <- TTP[!is.na(TTP$recurrence), ]
  TTP$time <- NA
  for (i in 1:nrow(TTP)) {
    if (TTP$recurrence[i] == 1) TTP$time[i] <- TTP$date_recur_1[i] - TTP$date_sx[i]
    if (TTP$recurrence[i] == 0) TTP$time[i] <- TTP$date_last_fu[i] - TTP$date_sx[i]
  }
  TTP <- TTP[!is.na(TTP$time), ]
  TTP
}

# ---- Overall Survival (OS): surgery date -> death, censored at last f/u ----
build_OS <- function(clinical, group_col = "chip_group") {
  OS <- select(clinical, study_id, chip_group = all_of(group_col), type_sx,
               date_sx, status, date_of_death, date_last_fu)

  OS$status[OS$status == 0] <- 2
  OS$status[OS$status == 1] <- 0
  OS$status[OS$status == 2] <- 1

  OS <- OS %>%
    mutate(date_sx = parse_date_flex(date_sx),
           date_of_death = parse_date_flex(date_of_death),
           date_last_fu = parse_date_flex(date_last_fu))

  OS <- OS[!is.na(OS$status), ]
  OS$time <- NA
  for (i in 1:nrow(OS)) {
    if (OS$status[i] == 1) OS$time[i] <- OS$date_of_death[i] - OS$date_sx[i]
    if (OS$status[i] == 0) OS$time[i] <- OS$date_last_fu[i] - OS$date_sx[i]
  }
  OS <- OS[!is.na(OS$time), ]
  OS <- filter(OS, time > 0)
  OS
}

# ---- Progression Free Survival (PFS): surgery date -> recurrence OR death ----
build_PFS <- function(clinical, group_col = "chip_group") {
  PFS <- select(clinical, study_id, chip_group = all_of(group_col), type_sx,
                date_sx, recurrence, status, date_recur_1, date_of_death, date_last_fu)

  PFS$recurrence[PFS$recurrence == 0] <- 2
  PFS$recurrence[PFS$recurrence == 1] <- 0
  PFS$recurrence[PFS$recurrence == 2] <- 1

  PFS$status[PFS$status == 0] <- 2
  PFS$status[PFS$status == 1] <- 0
  PFS$status[PFS$status == 2] <- 1

  PFS <- PFS %>%
    mutate(date_sx = parse_date_flex(date_sx),
           date_recur_1 = parse_date_flex(date_recur_1),
           date_of_death = parse_date_flex(date_of_death),
           date_last_fu = parse_date_flex(date_last_fu))

  PFS$progression <- 0
  PFS$date_progression <- ymd(NA)
  PFS$progression[PFS$recurrence == 1 | PFS$status == 1] <- 1
  PFS$date_progression[PFS$status == 1] <- PFS$date_of_death[PFS$status == 1]
  PFS$date_progression[PFS$recurrence == 1] <- PFS$date_recur_1[PFS$recurrence == 1]

  PFS$time <- NA
  for (i in 1:nrow(PFS)) {
    if (PFS$progression[i] == 1) PFS$time[i] <- PFS$date_progression[i] - PFS$date_sx[i]
    if (PFS$progression[i] == 0) PFS$time[i] <- PFS$date_last_fu[i] - PFS$date_sx[i]
  }
  PFS <- PFS[!is.na(PFS$time), ]
  PFS
}

# ---- Fibrosis: score taken as-is, levels 0-4 directly from the clinical data ----
build_fibrosis <- function(clinical, group_col = "chip_group") {
  fib <- select(clinical, study_id, chip_group = all_of(group_col), type_sx, fibrosis_score)
  fib <- filter(fib, !is.na(fibrosis_score))
  fib
}

TTP_all <- build_TTP(clinical)
OS_all <- build_OS(clinical)
PFS_all <- build_PFS(clinical)
Fibrosis_all <- build_fibrosis(clinical)
```

# Results

## All Patients

### Time to Progression

```{r ttp-all}
run_km_cox(TTP_all, "time", "recurrence", "chip_group",
           ylab = "Time to Progression", label = "TTP - All",
           outfile = "Plots/GeneVAF_Subgroups/TTP_All.pdf")
```

### Overall Survival

```{r os-all}
run_km_cox(OS_all, "time", "status", "chip_group",
           ylab = "Overall Survival", label = "OS - All",
           outfile = "Plots/GeneVAF_Subgroups/OS_All.pdf")
```

### Progression Free Survival

```{r pfs-all}
run_km_cox(PFS_all, "time", "progression", "chip_group",
           ylab = "Progression Free Survival", label = "PFS - All",
           outfile = "Plots/GeneVAF_Subgroups/PFS_All.pdf")
```

### Fibrosis

```{r fibrosis-all}
run_fibrosis(Fibrosis_all, "chip_group", label = "Fibrosis - All",
             table_outfile = "Plots/GeneVAF_Subgroups/Fibrosis_All.csv",
             plot_outfile = "Plots/GeneVAF_Subgroups/Fibrosis_All.pdf")
```

## Resection Only

```{r subset-resection}
TTP_resection <- TTP_all[TTP_all$type_sx == 0, ]
OS_resection <- OS_all[OS_all$type_sx == 0, ]
PFS_resection <- PFS_all[PFS_all$type_sx == 0, ]
Fibrosis_resection <- Fibrosis_all[Fibrosis_all$type_sx == 0, ]
```

### Time to Progression

```{r ttp-resection}
run_km_cox(TTP_resection, "time", "recurrence", "chip_group",
           ylab = "Time to Progression", label = "TTP - Resection",
           outfile = "Plots/GeneVAF_Subgroups/TTP_Resection.pdf")
```

### Overall Survival

```{r os-resection}
run_km_cox(OS_resection, "time", "status", "chip_group",
           ylab = "Overall Survival", label = "OS - Resection",
           outfile = "Plots/GeneVAF_Subgroups/OS_Resection.pdf")
```

### Progression Free Survival

```{r pfs-resection}
run_km_cox(PFS_resection, "time", "progression", "chip_group",
           ylab = "Progression Free Survival", label = "PFS - Resection",
           outfile = "Plots/GeneVAF_Subgroups/PFS_Resection.pdf")
```

### Fibrosis

```{r fibrosis-resection}
run_fibrosis(Fibrosis_resection, "chip_group", label = "Fibrosis - Resection",
             table_outfile = "Plots/GeneVAF_Subgroups/Fibrosis_Resection.csv",
             plot_outfile = "Plots/GeneVAF_Subgroups/Fibrosis_Resection.pdf")
```

## Transplant Only

```{r subset-transplant}
TTP_transplant <- TTP_all[TTP_all$type_sx == 1, ]
OS_transplant <- OS_all[OS_all$type_sx == 1, ]
PFS_transplant <- PFS_all[PFS_all$type_sx == 1, ]
Fibrosis_transplant <- Fibrosis_all[Fibrosis_all$type_sx == 1, ]
```

### Time to Progression

```{r ttp-transplant}
run_km_cox(TTP_transplant, "time", "recurrence", "chip_group",
           ylab = "Time to Progression", label = "TTP - Transplant",
           outfile = "Plots/GeneVAF_Subgroups/TTP_Transplant.pdf")
```

### Overall Survival

```{r os-transplant}
run_km_cox(OS_transplant, "time", "status", "chip_group",
           ylab = "Overall Survival", label = "OS - Transplant",
           outfile = "Plots/GeneVAF_Subgroups/OS_Transplant.pdf")
```

### Progression Free Survival

```{r pfs-transplant}
run_km_cox(PFS_transplant, "time", "progression", "chip_group",
           ylab = "Progression Free Survival", label = "PFS - Transplant",
           outfile = "Plots/GeneVAF_Subgroups/PFS_Transplant.pdf")
```

### Fibrosis

```{r fibrosis-transplant}
run_fibrosis(Fibrosis_transplant, "chip_group", label = "Fibrosis - Transplant",
             table_outfile = "Plots/GeneVAF_Subgroups/Fibrosis_Transplant.csv",
             plot_outfile = "Plots/GeneVAF_Subgroups/Fibrosis_Transplant.pdf")
```

# Results (Alternate CHIP Grouping)

This section repeats the exact same outcomes (TTP, OS, PFS, fibrosis) and
the same three cohorts (All, Resection, Transplant) as the Results section
above, but using the alternate 3-group classification defined earlier under
Alternate CHIP Grouping (`chip_group_alt`) instead of the original 4-group
one. Output files are written to a separate `AltGrouping` subfolder so the
original Results section's output files are left untouched.

```{r build-outcomes-alt}
dir.create("Plots/GeneVAF_Subgroups/AltGrouping", showWarnings = FALSE, recursive = TRUE)

TTP_alt <- build_TTP(clinical, group_col = "chip_group_alt")
OS_alt <- build_OS(clinical, group_col = "chip_group_alt")
PFS_alt <- build_PFS(clinical, group_col = "chip_group_alt")
Fibrosis_alt <- build_fibrosis(clinical, group_col = "chip_group_alt")
```

## All Patients

### Time to Progression

```{r ttp-all-alt}
run_km_cox(TTP_alt, "time", "recurrence", "chip_group",
           ylab = "Time to Progression", label = "TTP (Alt Grouping) - All",
           outfile = "Plots/GeneVAF_Subgroups/AltGrouping/TTP_All.pdf")
```

### Overall Survival

```{r os-all-alt}
run_km_cox(OS_alt, "time", "status", "chip_group",
           ylab = "Overall Survival", label = "OS (Alt Grouping) - All",
           outfile = "Plots/GeneVAF_Subgroups/AltGrouping/OS_All.pdf")
```

### Progression Free Survival

```{r pfs-all-alt}
run_km_cox(PFS_alt, "time", "progression", "chip_group",
           ylab = "Progression Free Survival", label = "PFS (Alt Grouping) - All",
           outfile = "Plots/GeneVAF_Subgroups/AltGrouping/PFS_All.pdf")
```

### Fibrosis

```{r fibrosis-all-alt}
run_fibrosis(Fibrosis_alt, "chip_group", label = "Fibrosis (Alt Grouping) - All",
             table_outfile = "Plots/GeneVAF_Subgroups/AltGrouping/Fibrosis_All.csv",
             plot_outfile = "Plots/GeneVAF_Subgroups/AltGrouping/Fibrosis_All.pdf")
```

## Resection Only

```{r subset-resection-alt}
TTP_resection_alt <- TTP_alt[TTP_alt$type_sx == 0, ]
OS_resection_alt <- OS_alt[OS_alt$type_sx == 0, ]
PFS_resection_alt <- PFS_alt[PFS_alt$type_sx == 0, ]
Fibrosis_resection_alt <- Fibrosis_alt[Fibrosis_alt$type_sx == 0, ]
```

### Time to Progression

```{r ttp-resection-alt}
run_km_cox(TTP_resection_alt, "time", "recurrence", "chip_group",
           ylab = "Time to Progression", label = "TTP (Alt Grouping) - Resection",
           outfile = "Plots/GeneVAF_Subgroups/AltGrouping/TTP_Resection.pdf")
```

### Overall Survival

```{r os-resection-alt}
run_km_cox(OS_resection_alt, "time", "status", "chip_group",
           ylab = "Overall Survival", label = "OS (Alt Grouping) - Resection",
           outfile = "Plots/GeneVAF_Subgroups/AltGrouping/OS_Resection.pdf")
```

### Progression Free Survival

```{r pfs-resection-alt}
run_km_cox(PFS_resection_alt, "time", "progression", "chip_group",
           ylab = "Progression Free Survival", label = "PFS (Alt Grouping) - Resection",
           outfile = "Plots/GeneVAF_Subgroups/AltGrouping/PFS_Resection.pdf")
```

### Fibrosis

```{r fibrosis-resection-alt}
run_fibrosis(Fibrosis_resection_alt, "chip_group", label = "Fibrosis (Alt Grouping) - Resection",
             table_outfile = "Plots/GeneVAF_Subgroups/AltGrouping/Fibrosis_Resection.csv",
             plot_outfile = "Plots/GeneVAF_Subgroups/AltGrouping/Fibrosis_Resection.pdf")
```

## Transplant Only

```{r subset-transplant-alt}
TTP_transplant_alt <- TTP_alt[TTP_alt$type_sx == 1, ]
OS_transplant_alt <- OS_alt[OS_alt$type_sx == 1, ]
PFS_transplant_alt <- PFS_alt[PFS_alt$type_sx == 1, ]
Fibrosis_transplant_alt <- Fibrosis_alt[Fibrosis_alt$type_sx == 1, ]
```

### Time to Progression

```{r ttp-transplant-alt}
run_km_cox(TTP_transplant_alt, "time", "recurrence", "chip_group",
           ylab = "Time to Progression", label = "TTP (Alt Grouping) - Transplant",
           outfile = "Plots/GeneVAF_Subgroups/AltGrouping/TTP_Transplant.pdf")
```

### Overall Survival

```{r os-transplant-alt}
run_km_cox(OS_transplant_alt, "time", "status", "chip_group",
           ylab = "Overall Survival", label = "OS (Alt Grouping) - Transplant",
           outfile = "Plots/GeneVAF_Subgroups/AltGrouping/OS_Transplant.pdf")
```

### Progression Free Survival

```{r pfs-transplant-alt}
run_km_cox(PFS_transplant_alt, "time", "progression", "chip_group",
           ylab = "Progression Free Survival", label = "PFS (Alt Grouping) - Transplant",
           outfile = "Plots/GeneVAF_Subgroups/AltGrouping/PFS_Transplant.pdf")
```

### Fibrosis

```{r fibrosis-transplant-alt}
run_fibrosis(Fibrosis_transplant_alt, "chip_group", label = "Fibrosis (Alt Grouping) - Transplant",
             table_outfile = "Plots/GeneVAF_Subgroups/AltGrouping/Fibrosis_Transplant.csv",
             plot_outfile = "Plots/GeneVAF_Subgroups/AltGrouping/Fibrosis_Transplant.pdf")
```
